A Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-02 02-ceu.tex
Revisão Seção 2.2

\subsection{Shared-Memory Concurrency}
    - referenciar warnings

@@

\begin{figure}[h]
\begin{minipage}[h]{0.45\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3em]
input void A;
input void B;
var int x = 1;
par/and do
    await A;
    x = x + 1;
with
    await B;
    x = x * 2;
end
\end{lstlisting}
\centering\small{\ax Accesses to \code{x} are never concurrent.}
\end{minipage}
%
\begin{minipage}[h]{0.45\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3em]
input void A;
var int y = 1;
par/and do
    await A;
    y = y + 1;
with
    await A;
    y = y * 2;
end
\end{lstlisting}
\centering\small{\bx Accesses to \code{y} are concurrent but deterministic.}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Shared-memory concurrency in \CEU:
example \ax is safe because the trails access \code{x} atomically in different
reactions;
example \bx is unsafe because both trails access \code{y} in the same reaction.
\label{lst.shared}
}
\end{figure}

In \CEU, when multiple trails are active during the same reaction, they are
scheduled in lexical order, i.e., in the order they appear in the program
source code.
%
For instance, consider the two examples in Figure~\ref{lst.shared}, both
defining a shared variable (ln. 2), and assigning to it in parallel trails (ln.
5, 8).

@@

In the example \bx, the two assignments to \code{y} are simultaneous because
they execute in reaction to the same event \code{A}.
Since \CEU employs lexical order for intra-reaction statements, the execution
is still deterministic, and \code{y} always becomes \code{4} (\code{(1+1)*2}).
%
However, that an (apparently innocuous) change in the order of trails modifies
the behavior of the program.
%
To mitigate this threat, \CEU performs concurrency checks at compile time to
detect conflicting accesses to shared variables:
if a variable is written in a trail segment, then a concurrent trail segment
cannot read or write to that variable~\cite{ceu.sensys13}.
%
Nonetheless, the static checks are optional and do not affect the semantics of
the language.

>>>>

\subsection{Shared-Memory Concurrency}

@@

\begin{figure}[h]
\begin{minipage}[h]{0.45\linewidth}
\begin{lstlisting}[numbers=right]
input void A;
input void B;
var int x = 1;
par/and do
    await A;
    x = x + 1;
with
    await B;
    x = x * 2;
end
\end{lstlisting}
\centering\small{\ax Accesses to \code{x} are never concurrent.}
\end{minipage}
%
\begin{minipage}[h]{0.53\linewidth}
%\begin{lstlisting}
\begin{lstlisting}[xleftmargin=2em]
input void A;
//
var int y = 1;
par/and do
    await A;
    y = y + 1;
with
    await A;
    y = y * 2;
end

\end{lstlisting}
\centering\small{\bx Accesses to \code{y} are concurrent but deterministic.}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Shared-memory concurrency in \CEU:
example \ax is safe because the trails access \code{x} atomically in different
reactions;
example \bx is unsafe because both trails access \code{y} in the same reaction.
\label{lst.shared}
}
\end{figure}

In \CEU, when multiple trails are active during the same reaction, they are
scheduled in lexical order, i.e., in the order they appear in the program
source code.
%
For instance, consider the two examples in Figure~\ref{lst.shared}, both
defining shared variables (ln. 3), and assigning to them in parallel trails
(ln. 6, 9).

@@

In the example \bx, the two assignments to \code{y} are simultaneous because
they execute in reaction to the same event \code{A}.
Since \CEU employs lexical order for intra-reaction statements, the execution
is still deterministic, and \code{y} always becomes \code{4} (\code{(1+1)*2}).
%
However, note that an apparently innocuous change in the order of trails
modifies the behavior of the program.
%
To mitigate this threat, \CEU performs concurrency checks at compile time to
detect conflicting accesses to shared variables:
if a variable is written in a trail segment, then a concurrent trail segment
cannot read or write to that variable~\cite{ceu.sensys13}.
%
Nonetheless, the static checks are optional and do not affect the semantics of
the language.
